---
title: "NMF"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading packages and data
NMF was performed on 


```{r}
library(data.table)
library(readxl)
library(NMF)
library(umap)
library(pheatmap)
library(fossil)
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
library(RColorBrewer)
library(ggplot2)
library(peakSeason)

```

## Optimal number of components
Cophenetic correlation is a measure of goodness of fit.

```{r echo=FALSE}
nmf_trial_nrows <- readRDS("/omics/groups/OE0219/internal/AnnaR/NMF/nmf_trial_nrows.RDS")
trial_nrows <- c(5000, 10000, 15000, 20000, 25000, 30000, 35000, 35838)


row_cols = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[2:9]
par(mar = c(4, 5, 1, 1), cex = 0.8)

plot(nmf_trial_nrows[[1]]$cor_metric$rank, nmf_trial_nrows[[1]]$cor_metric$cophenetic , axes = FALSE, ylim = c(0.85, 1), xlim = c(2, 10), pch = 16, col = row_cols[1], xlab = NA, ylab = NA)

abline(h = seq(0.85, 1, 0.01), v = 2:10, col = grDevices::adjustcolor(col = 'gray70', alpha.f = 0.7), lty = 2)

lines(nmf_trial_nrows[[1]]$cor_metric$rank, nmf_trial_nrows[[1]]$cor_metric$cophenetic, lwd = 1.2, col = row_cols[1])
points(nmf_trial_nrows[[1]]$cor_metric$rank, nmf_trial_nrows[[1]]$cor_metric$cophenetic, pch = 19, col = row_cols[1])

for(i in 2:length(nmf_trial_nrows)){
  lines(x = nmf_trial_nrows[[i]]$cor_metric$rank, nmf_trial_nrows[[i]]$cor_metric$cophenetic, col = row_cols[i])
  points(nmf_trial_nrows[[i]]$cor_metric$rank, nmf_trial_nrows[[i]]$cor_metric$cophenetic, pch = 19, col = row_cols[i])
}

axis(side = 1, at = 2:10, labels = 2:10, font = 1)
axis(side = 2, at = seq(0.85, 1, 0.01), font = 1, las = 2, cex = 1.4)

mtext(text = "Number of components", side = 1, line = 2.5, adj = 0, cex = 0.8)
mtext(text = "Cophenetic-correlation metric", side = 2, line = 3.5, adj = 0, cex = 0.8)

legend(x = 9, y = 1.0, legend = trial_nrows, col = row_cols, pch = 19, cex = 0.8)

```
Optimal number of components seems to be n= x

## Cluster stability
Re-run nmf at n=3 for above range of probes to access cluster stability.

## Rand-Index to access cluster integrity 
Check how stable the clusters are by estimating Rand-index between all k-means clustering results. Performed k-means clustering on canepi-srv1.

```{r}
#load clustering results
nmf_k3_clusters = readRDS("/omics/groups/OE0219/internal/AnnaR/NMF/nmf_k3_clusters.RDS")
 
#pairwise rand-index estimation
interactions = sapply(1:ncol(nmf_k3_clusters), function(i) {
  sapply(1:ncol(nmf_k3_clusters), function(j) {
    fossil::rand.index(group1 = nmf_k3_clusters[, i], group2 = nmf_k3_clusters[, j])
  })
})

#Draw rand-index
rownames(interactions) = colnames(interactions) = colnames(nmf_k3_clusters)

interactions[lower.tri(x = interactions)] = NA
m <- nrow(interactions)
n <- ncol(interactions)

par(bty="n", mar = c(1, 4, 2, 5)+.1, las=2, tcl=-.33, cex = 0.6)
image(
  x = 1:n,
  y = 1:m,
  interactions,
  col = rev(RColorBrewer::brewer.pal(9, "PRGn")),
  xaxt = "n",
  yaxt = "n",
  xlab = "",
  ylab = "",
  xlim = c(0, n+2),
  ylim = c(0, n+2)
)
abline(h=0:n+.5, col="white", lwd=.5)
abline(v=0:n+.5, col="white", lwd=.5)

image(y = seq(0.5*nrow(interactions), 0.9*nrow(interactions), length.out = 8), x=rep(n,2)+c(2,2.5)+1, z=matrix(c(1:8), nrow=1), col = rev(RColorBrewer::brewer.pal(8,"PRGn")), add=TRUE)
atLims = seq(0.5*nrow(interactions), 0.9*nrow(interactions), length.out = 7)
atLimsLabs = round(seq(range(interactions, na.rm = T)[1], range(interactions, na.rm = T)[2], length.out = 7), 2)
axis(side = 4, at = atLims,  tcl = -.15, labels = atLimsLabs, las = 1, lwd = .5)
mtext(side = 4, at = median(atLims), "Rand Index", las = 3, cex = 0.9, line = 3, font = 2)

mtext(side = 2, at = 1:m, text = gsub(pattern = "nprobes_", replacement = "", x = colnames(interactions)), cex = 0.7, font = 1)
mtext(side = 3, at = 1:n, text = gsub(pattern = "nprobes_", replacement = "", x = colnames(interactions)), las = 2, cex = 0.7, font = 1, line = -2) 

#

```
## Correlation matrix
Simple visualization to see movement of samples for different clustering results
